The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-08-31 paper/paper_word_output.Rmd
Started word output Rmd

---
title: The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

blinded: 0

authors: 
- name: Albert Y. Kim 
  affiliation: Program in Statistical & Data Sciences, Smith College
  thanks: "Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."
- name: David N. Allen
  affiliation: Biology Department, Middlebury College
- name: Simon P. Couch
  affiliation: Mathematics Department, Reed College

abstract: 
- |
  | 1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
  | 2. We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.  Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
  | 3. We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites. Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind. We highlight the importance of spatial cross-validation when interpreting model results. 
  | 4. The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model. These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data. Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.

keywords:
- forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation

bibliography: paper.bib
output:
  rticles::asa_article
link-citations: yes
header-includes:
- \usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace, fancyhdr}
# Add line numbers:
- \linenumbers
# Add running header:
- \pagestyle{fancy}
- \fancyhead[LO,LE]{forestecology R package}
- \renewcommand{\sectionmark}[1]{\markright{#1}{}}
---

@@

@@

# Acknowledgments

We thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.

# Author's contributions

AYK and DNA conceived the ideas and coded a draft of the package. AYK wrote an initial manuscript draft. SPC rewrote much of the package's code to align with R and "tidy" best practices [@wickham_welcome_2019]. All authors contributed to subsequent drafts and gave final approval for manuscript.

# Data accessibility

We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and on Zenodo upon acceptance. The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

>>>>

---
title: The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees
authors: 
- name: Albert Y. Kim 
  affiliation: Program in Statistical & Data Sciences, Smith College
  thanks: "Assistant Professor, Statistical & Data Sciences, Smith College, Northampton, MA 01063 (e-mail: akim04@smith.edu)."
- name: David N. Allen
  affiliation: Biology Department, Middlebury College
- name: Simon P. Couch
  affiliation: Mathematics Department, Reed College
bibliography: paper.bib
output:
  bookdown::word_document2:
  fig_caption: yes
link-citations: yes
header-includes:
- \usepackage{xcolor, soul, xspace, float, subfig, lineno, setspace, fancyhdr}
# Add line numbers:
# - \linenumbers
# Add running header:
- \pagestyle{fancy}
- \fancyhead[LO,LE]{forestecology R package}
- \renewcommand{\sectionmark}[1]{\markright{#1}{}}
---

@@

# Abstract

1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
2. We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.  Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
3. We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites. Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind. We highlight the importance of spatial cross-validation when interpreting model results. 
4. The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model. These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data. Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.

Keywords: forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation

@@

# Acknowledgements {-}

We thank Sophie Li for their feedback on the package interface.

# Conflict of Interest {-}

We declare no competing interests.

# Author's contributions {-}

**Albert Y. Kim**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (lead); Writing-review & editing (equal). **David N. Allen**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (supporting); Writing-review & editing (equal). **Simon P. Couch**: Software (supporting); Writing-review & editing (supporting).

# Data accessibility {-}

All data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and are archived on Zenodo at https://doi.org/10.5281/zenodo.5348060. The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].
